Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software

### 2022-01-21 manuscript/manuscript-speed-accuracy.tex
Removed a benchmark with a conflicted author, updated figures, manuscript etc.

\usepackage[utf8]{inputenc} %unicode support
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[hyphens]{url}
%\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{todonotes}
%\usepackage{maori}
%\graphicspath{{../figures/}}
\usepackage{xr}
%\externaldocument{supplementary}

@@

%%% Put your definitions there:
\startlocaldefs
\def\numTools{499}
%cat meanRankSpeedData.tsv | cut -f 4 | grep -v ^method$ | wc -l
\def\numBenchmarkPubs{69}
%cat speed-vs-accuracy-toolRanks2005-2020.tsv | cut -f 1 | sort  | uniq | grep -E '[0-9]'  | wc -l
\def\numBenchmarks{134}
%cat rawRankSpeedData2005-2020.tsv | cut -f 1 | grep -v ^testId$ | uniq -c | wc -l
\endlocaldefs

@@

%\Abstract{
\begin{abstractbox}
  \begin{abstract} % abstract
    \parttitle{Background} Computational biology provides widely used and
    powerful software tools for testing and making inferences about
    biological data. In the face of rapidly increasing volumes of
    data, heuristic methods that trade software speed for accuracy may
    be employed.  We are have studied these trade-offs using the
    results of a large number of independent software benchmarks, and
    evaluated whether external factors are indicative of accurate
    software.

    \parttitle{Method} We have extracted accuracy and speed ranks from
    independent benchmarks of different bioinformatic software tools,
    and evaluated whether the speed, author reputation, journal
    impact, recency and developer efforts are indicative of accuracy.

    \parttitle{Results} We found that software speed, author reputation,
    journal impact, number of citations and age are all unreliable
    predictors of software accuracy. This is unfortunate because
    citations, author and journal reputation are frequently cited
    reasons for selecting software tools. However, GitHub-derived
    records and high version numbers show that the accurate
    bioinformatic software tools are generally the product of many
    improvements over time, often from multiple developers.

    \parttitle{Discussion} We also find that the field of bioinformatics has a
    large excess of slow and inaccurate software tools, and this is
    consistent across many sub-disciplines. Meanwhile, there are few
    tools that are middle-of-road in terms of accuracy and speed
    trade-offs. We hypothesise that a form of publication-bias
    influences the publication and development of bioinformatic
    software. In other words, software that is intermediate in terms
    of both speed and accuracy may be difficult to publish - possibly
    due to author, editor and reviewer practices. This leaves an
    unfortunate hole in the literature as the ideal tools may fall
    into this gap. For example, high accuracy tools are not always
    useful if years of CPU time are required, while high speed is not
    useful if the results are also inaccurate.
%}
\end{abstract}

@@

Data on the number of updates to software tools from GitHub such as
the version number, {\color{black}and numbers of contributors, commits, forks and issues } was significantly
correlated with software accuracy (respective Spearman’s rhos =
{\color{black}0.15, 0.21, 0.22, 0.23, 0.23 and respective Benjamini \& Hochberg
  corrected P-values = $5.4\times 10^{-4}$, $1.1\times 10^{-3}$, $7.9\times 10^{-4}$, $3.4\times 10^{-4}$, $3.2\times 10^{-4}$, Additional file 1: Figures S6}). %See arrays "spearmansA" and "spearmansAP" in prettyPlot.R
The significance of these features was
further confirmed with a permutation test
(Figure~\ref{fig:allfactors}B).  These features were not correlated
with speed however (see Figure~\ref{fig:allfactors}A {\color{black}\& Additional file 1: Figures S5\&S6}).
We also found that
reputation metrics such as citations, author and journal H-indices,
and the age of tools were generally \textbf{not} correlated with
either tool accuracy or speed (Figure~\ref{fig:allfactors}A\&B).

@@

In order to gain a deeper understanding of the distribution of
available bioinformatic software tools on a speed versus accuracy
landscape, we ran a permutation test. The ranks extracted
from each benchmark were randomly permuted, generating 1,000
randomized speed and accuracy ranks. In the cells of a $3\times3$
grid spanning the normalised speed and accuracy ranks we computed a
Z-score for the observed number of tools in a cell, compared to the
expected distributions generated by 1,000 randomized ranks. The results
of this are shown in Figure~\ref{fig:speedaccuracy}. We identified 4
of 9 bins where there was a significant excess or dearth of tools. For
example, there was an excess of ``slow and inaccurate'' software (Z=3.39,
P-value=$3.5\times 10^{-4}$), with more moderate excess of ``slow and accurate'' and ``fast and accurate'' software (Z=2.49 \& 1.7, P=$6.3\times 10^{-3}$ \& 0.04 respectively). We find that only the ``fast and inaccurate'' extreme class is at approximately the
expected proportions based upon the permutation test (Figure~\ref{fig:speedaccuracy}B).

@@

\section*{Availability of data and materials}
%\section*{Data availability}
Raw datasets, software and documents are available under a CC-BY license:\\
%\fussy
%\url{https://docs.google.com/spreadsheets/d/14xIY2PHNvxmV9MQLpbzSfFkuy1RlzDHbBOCZLJKcGu8/edit?usp=sharing}\\
%and here:\\
%\fussy
%\url{https://dx.doi.org/10.6084/m9.figshare.4299320.v1}\\
%\sloppy
%Additional documentation, code, figures and raw data is available here:\\
\fussy
\url{https://github.com/Gardner-BinfLab/speed-vs-accuracy-meta-analysis}
\sloppy

>>>>

\usepackage[utf8]{inputenc} %unicode support
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[hyphens]{url}
%\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{todonotes}
%\usepackage{maori}
%\graphicspath{{../figures/}}
\usepackage{xr}
%\externaldocument{supplementary}
\usepackage{url}
\usepackage{hyperref}

@@

%%% Put your definitions there:
\startlocaldefs
\def\numTools{498}
%cat meanRankSpeedData.tsv | cut -f 4 | grep -v ^method$ | wc -l
\def\numBenchmarkPubs{68}
%cat speed-vs-accuracy-toolRanks2005-2020.tsv | cut -f 1 | sort  | uniq | grep -E '[0-9]'  | wc -l
\def\numBenchmarks{133}
%cat rawRankSpeedData2005-2020.tsv | cut -f 1 | grep -v ^testId$ | uniq -c | wc -l
\endlocaldefs

@@

%\Abstract{
\begin{abstractbox}
  \begin{abstract} % abstract

    \parttitle{Background:} Computational biology provides software
    tools for testing and making inferences about biological data. In
    the face of increasing volumes of data, heuristic methods
    that trade software speed for accuracy may be employed. We
    have studied these trade-offs using the results of a large number
    of independent software benchmarks, and evaluated whether external
    factors, including speed, author reputation, journal impact,
    recency and developer efforts, are indicative of accurate
    software.

    \parttitle{Results:} We find that software speed, author
    reputation, journal impact, number of citations and age are
    unreliable predictors of software accuracy. This is unfortunate
    because these are frequently
    cited reasons for selecting software tools. However,
    GitHub-derived statistics and high version numbers show that accurate
    bioinformatic software tools are generally the product of many
    improvements over time. We also
    find an excess of slow and inaccurate bioinformatic software
    tools, and this is consistent across many
    sub-disciplines. Meanwhile, there are few tools that are
    middle-of-road in terms of accuracy and speed trade-offs.

    \parttitle{Conclusions:} Our findings indicate that accurate
    bioinformatic software is primarily the product of long-term
    commitments to software-development. In addition, we hypothesise that
    bioinformatics software suffers from publication bias. Software
    that is intermediate in terms of both speed and accuracy may be
    difficult to publish - possibly due to author, editor and reviewer
    practises. This leaves an unfortunate hole in the literature, as
    ideal tools may fall into this gap. High accuracy tools are
    not always useful if they are slow, while high speed is not useful
    if the results are also inaccurate.

%}
\end{abstract}

@@

Data on the number of updates to software tools from GitHub such as
the version number, {\color{black}and numbers of contributors, commits, forks and issues } was significantly
correlated with software accuracy (respective Spearman’s rhos =
{\color{black}0.15, 0.21, 0.22, 0.23, 0.23 and respective Benjamini \& Hochberg
  corrected P-values = $6.7\times 10^{-4}$, $1.1\times 10^{-3}$, $8.4\times 10^{-4}$, $3.4\times 10^{-4}$, $3.1\times 10^{-4}$, Additional file 1: Figures S6}). %See arrays "spearmansA" and "spearmansAP" in prettyPlot.R
The significance of these features was
further confirmed with a permutation test
(Figure~\ref{fig:allfactors}B).  These features were not correlated
with speed however (see Figure~\ref{fig:allfactors}A {\color{black}\& Additional file 1: Figures S5\&S6}).
We also found that
reputation metrics such as citations, author and journal H-indices,
and the age of tools were generally \textbf{not} correlated with
either tool accuracy or speed (Figure~\ref{fig:allfactors}A\&B).

@@

In order to gain a deeper understanding of the distribution of
available bioinformatic software tools on a speed versus accuracy
landscape, we ran a permutation test. The ranks extracted
from each benchmark were randomly permuted, generating 1,000
randomized speed and accuracy ranks. In the cells of a $3\times3$
grid spanning the normalised speed and accuracy ranks we computed a
Z-score for the observed number of tools in a cell, compared to the
expected distributions generated by 1,000 randomized ranks. The results
of this are shown in Figure~\ref{fig:speedaccuracy}. We identified 4
of 9 bins where there was a significant excess or dearth of tools. For
example, there was an excess of ``slow and inaccurate'' software ({\color{black}Z=3.40,
P-value=$3.3\times 10^{-4}$}), with more moderate excess of ``slow and accurate'' and ``fast and accurate'' software (Z=2.49 \& 1.7, P=$6.3\times 10^{-3}$ \& 0.04 respectively). We find that only the ``fast and inaccurate'' extreme class is at approximately the
expected proportions based upon the permutation test (Figure~\ref{fig:speedaccuracy}B).

%#Z=
%print(zScores[1,1])
%#P=
%pnorm(zScores[1,1], mean = 0, sd = 1, lower.tail = F)

@@

\section*{Availability of data and materials}
%\section*{Data availability}
Raw datasets, software and documents are available under a CC-BY license at Github \cite{Gardner2022a} and FigShare \cite{Gardner2022b}.
%\fussy
%\url{https://docs.google.com/spreadsheets/d/14xIY2PHNvxmV9MQLpbzSfFkuy1RlzDHbBOCZLJKcGu8/edit?usp=sharing}\\
%and here:\\
%\fussy
%\url{https://dx.doi.org/10.6084/m9.figshare.4299320.v1}\\
%\sloppy
%Additional documentation, code, figures and raw data is available here:\\
%\fussy
%Github: \url{https://github.com/Gardner-BinfLab/speed-vs-accuracy-meta-analysis}\\
%FigShare: \url{https://doi.org/10.6084/m9.figshare.15121818.v2}

\sloppy
